\begin{thebibliography}{10}
\bibliographystyle{aaai}


\bibitem{Abiteboul95}
S.~Abiteboul, R.~Hull and V.~Vianu.
\newblock {\em Foundations of Databases}.
\newblock Addison-Wesley,   1995.

\bibitem{Rubin1983b}
Rosenbaum, P. R. and Rubin, D. B.
\newblock
The Central Role of
the Propensity Score in Observational Studies for Causal Effects.
\newblock {\em Biometrika}, 1983, 70(1):41-55.


\bibitem{Stuart10}
 E.~A.~Stuart.
\newblock
Matching Methods for Causal Inference: A Review and a Look Forward.
\newblock {\em journal of the Institute of Mathematical Statistics}, 2010, 25(1):1.


\bibitem{IacKinPor09}
Iacus, S. M., Gary K. and Giuseppe P.
\newblock
Causal Inference Without Balance Checking: Coarsened Exact Matching
\newblock {\em Political Analysis}, 2011.


\bibitem{FAA08}
FAA. Pilot’s Handbook of Aeronautical Knowledge.
\newblock  United States Department of Transportation, Federal Aviation Administration, Airman
Testing Standards Branch, AFS-630, P.O. Box 25082, Oklahoma City, OK 73125. 2008.


\bibitem{weba1}
\newblock   Factors Affecting Aircraft Performance During Takeoff and Climb.
\newblock  \href{http://www.skybrary.aero/index.php/AP4ATCO\_-\_Factors\_Affecting\_Aircraft\_Performance\_During\_Takeoff\_and\_Climb}
{\small http://www.skybrary.aero/index.php/AP4ATCO\_-\_Factors\_Affecting\_Aircraft\_Performance\_During\_Takeoff\_and\_Climb}


\bibitem{weba2}
\newblock  High and Low Pressure.
\newblock  \href{http://www.weatherworksinc.com/high-low-pressure}{http://www.weatherworksinc.com/high-low-pressure}



\bibitem{Avis83}
D.~Avis.
\newblock A Survey of Heuristics for the Weighted
Matching Problem.
\newblock {\em Journal of the ACM (JACM)}, 1983, 35(4): 769-776.


\bibitem{Ball10}
M.~Ball et. al.
\newblock Total delay impact
study.
\newblock NEXTOR Research Symposium, Washington DC, January, 2010.







\bibitem{DBLP:conf/pods/BeameKS14}
P.~Beame, P.~Koutris, and D.~Suciu.
\newblock Skew in parallel query processing.
\newblock In {\em Proceedings of the 33rd {ACM} {SIGMOD-SIGACT-SIGART}
  Symposium on Principles of Database Systems, PODS'14, Snowbird, UT, USA, June
  22-27, 2014}, pages 212--223, 2014.

\bibitem{DBLP:journals/debu/BhattacharyaG06}
I.~Bhattacharya and L.~Getoor.
\newblock Collective entity resolution in relational data.
\newblock {\em {IEEE} Data Eng. Bull.}, 29(2):4--12, 2006.

\bibitem{DBLP:journals/tkdd/BhattacharyaG07}
I.~Bhattacharya and L.~Getoor.
\newblock Collective entity resolution in relational data.
\newblock {\em {TKDD}}, 1(1), 2007.

\bibitem{DBLP:conf/sigmod/ChaudhuriMN99}
S.~Chaudhuri, R.~Motwani, and V.~R. Narasayya.
\newblock On random sampling over joins.
\newblock In {\em {ACM} {SIGMOD} International Conference on Management of
  Data, SIGMOD'99}, pages 263--274, 1999.

\bibitem{DBLP:journals/ftdb/CheneyCT09}
J.~Cheney, L.~Chiticariu, and W.-C. Tan.
\newblock Provenance in databases: Why, how, and where.
\newblock {\em Foundations and Trends in Databases}, 1(4):379--474, 2009.

\bibitem{ChocklerH04}
H.~Chockler and J.~Y. Halpern.
\newblock Responsibility and blame: A structural-model approach.
\newblock {\em J. Artif. Intell. Res. (JAIR)}, 22:93--115, 2004.

Cochran, William G. "The effectiveness of adjustment by subclassification in removing bias in observational studies." Biometrics (1968): 295-313.

\bibitem{Cochran1968}
W.~G. Cochran.
\newblock The effectiveness of adjustment by subclassification in removing bias in observational studies.
\newblock {\em {Biometrics}},  1968, pp. 295-313.

\bibitem{Cochran1953}
W.~G. Cochran.
\newblock Matching in analytical studies.
\newblock {\em {American Journal of Public Health}}, 43:684--691, 1953.


\bibitem{DBLP:journals/ftdb/CormodeGHJ12}
G.~Cormode, M.~N. Garofalakis, P.~J. Haas, and C.~Jermaine.
\newblock Synopses for massive data: Samples, histograms, wavelets, sketches.
\newblock {\em Foundations and Trends in Databases}, 4(1-3):1--294, 2012.

\bibitem{DBLP:journals/tods/CuiWW00}
Y.~Cui, J.~Widom, and J.~L. Wiener.
\newblock Tracing the lineage of view data in a warehousing environment.
\newblock {\em {ACM} Trans. Database Syst.}, 25(2):179--227, 2000.

\bibitem{DBLP:conf/icdt/DarwenDF12}
H.~Darwen, C.~J. Date, and R.~Fagin.
\newblock A normal form for preventing redundant tuples in relational
  databases.
\newblock In {\em 15th International Conference on Database Theory, {ICDT} '12,
  Berlin, Germany, March 26-29, 2012}, pages 114--126, 2012.

\bibitem{flightdata}
{Dataset: Airline On-Time Performance}.
\newblock {\mbox{\url{http://www.transtats.bts.gov/}}}.
\newblock {\em {Bureau of Transportation Statistics, United States Department
  of Transportation}}.

\bibitem{dblp}
{Dataset: DBLP}.
\newblock \mbox{\url{http://dblp.uni-trier.de}}.
\newblock {\em {DBLP - Computer Science Bibliography}}.

\bibitem{natality}
{Dataset: Natality}.
\newblock \mbox{\url{http://www.cdc.gov/nchs/data_access/ftp_data.htm}}.
\newblock {\em {Birth Records in the United States in 2010 from Centers for
  Disease Control and Prevention (CDC) and National Center for Health
  Statistics (NCHS)}}.

\bibitem{nsf-awards}
{Dataset: NSF Awards}.
\newblock \mbox{\url{http://www.nsf.gov/awardsearch/download.jsp}}.
\newblock {\em {National Science Foundation}}.


\bibitem{Richardson11}
 X.~De~Luna,I.~Waernbaum and T.S.~Richardson,
\newblock {Covariate selection for the nonparametric estimation of an average treatment effect. }.
\newblock {\em Biometrika}, volume~26, pages 1058--1063, 19


De Luna, X., Waernbaum, I. and Richardson, T.S., 2011. Covariate selection for the nonparametric estimation of an average treatment effect. Biometrika, p.asr041.

\bibitem{deraedt:ml96}
L.~De~Raedt and L.~Dehaspe.
\newblock {Clausal Discovery}.
\newblock In {\em Machine Learning}, volume~26, pages 1058--1063, 1996.

\bibitem{DBLP:conf/kr/BroeckMD14}
G.~V. den Broeck, W.~Meert, and A.~Darwiche.
\newblock Skolemization for weighted first-order model counting.
\newblock In {\em Principles of Knowledge Representation and Reasoning:
  Proceedings of the Fourteenth International Conference, {KR} 2014, Vienna,
  Austria, July 20-24, 2014}, 2014.

\bibitem{deshpande:mmudchapter09}
A.~Deshpande, L.~Getoor, and P.~Sen.
\newblock Graphical models for uncertain data.
\newblock In C.~Aggarwal, editor, {\em Managing and Mining Uncertain Data}.
  Springer, 2009.

\bibitem{dehejia99}
R.~H.~Wehejia and S.~Wahba
\newblock Causal effects in nonexperimental
studies: Re-evaluating the evaluation of training
programs.
\newblock {\em J. Amer. Statist. Assoc}.
  Springer,1999, 94:1053–1062.




\bibitem{DBLP:conf/sigmod/Fagin79}
R.~Fagin.
\newblock Normal forms and relational database operators.
\newblock In {\em {ACM} {SIGMOD} International Conference on Management of
  Data, SIGMOD'79}, pages 153--160, 1979.

\bibitem{Fisher1926}
R.~Fisher.
\newblock The arrangement of field experiments.
\newblock In S.~Kotz and N.~Johnson, editors, {\em Breakthroughs in
  Statistics}, Springer Series in Statistics, pages 82--91. Springer New York,
  1992.

\bibitem{Fisher1935design}
R.~A. Fisher.
\newblock {\em The design of experiments}.
\newblock Oliver and Boyd, Oxford, England, 1935.

\bibitem{fowler:apr11}
J.~Fowler, M.~Heaney, D.~Nickerson, J.~Padgett, and B.~Sinclair.
\newblock Causality in politcal networks.
\newblock {\em American Politics Research}, 39(2), 2011.

\bibitem{friedman:ijcai99}
N.~Friedman, L.~Getoor, D.~Koller, and A.~Pfeffer.
\newblock Learning probabilistic relational models.
\newblock In {\em Proceedings of the 16th International Joint Conference on
  Artificial Intelligence - Volume 2}, IJCAI'99, pages 1300--1307, 1999.

\bibitem{DBLP:journals/amai/GeigerP90}
D.~Geiger and J.~Pearl.
\newblock Logical and algorithmic properties of independence and their
  application to bayesian networks.
\newblock {\em Ann. Math. Artif. Intell.}, 2:165--178, 1990.

\bibitem{getoor:jmlr02}
L.~Getoor, N.~Friedman, D.~Koller, and B.~Taskar.
\newblock Learning probabilistic models of link structure.
\newblock {\em Journal of Machine Learning Research}, 3:679--707, 2002.

\bibitem{getoor:kdd13-tutorial}
L.~Getoor and A.~Machanavajjhala.
\newblock Tutorial on entity resolution in big data.
\newblock In {\em ACM SIGKDD International Conference on Knowledge Discovery
  and Data Mining, KDD'13}, page 1527, 2013.

\bibitem{lise:ben:book:2007}
L.~Getoor and B.~Taskar.
\newblock {\em Introduction to Statistical Relational Learning}.
\newblock MIT Press, 2007.

\bibitem{GKT07-semirings}
T.~J. Green, G.~Karvounarakis, and V.~Tannen.
\newblock Provenance semirings.
\newblock In {\em Proceedings of the Twenty-sixth ACM SIGMOD-SIGACT-SIGART
  Symposium on Principles of Database Systems}, PODS '07, pages 31--40, 2007.

\bibitem{griffin1995incremental}
T.~Griffin and L.~Libkin.
\newblock Incremental maintenance of views with duplicates.
\newblock In {\em {ACM} {SIGMOD} International Conference on Management of
  Data, SIGMOD'95}, pages 328--339, 1995.

\bibitem{grohe-book}
M.~Grohe.
\newblock Descriptive complexity, canonisation, and definable graph structure
  theory, 2013.
\newblock Book (Manuscript).

\bibitem{DBLP:journals/ipl/GyssensNG14}
M.~Gyssens, M.~Niepert, and D.~V. Gucht.
\newblock On the completeness of the semigraphoid axioms for deriving arbitrary
  from saturated conditional independence statements.
\newblock {\em Inf. Process. Lett.}, 114(11):628--633, 2014.

\bibitem{HalpernP01}
J.~Y. Halpern and J.~Pearl.
\newblock Causes and explanations: {A} structural-model approach: Part 1:
  Causes.
\newblock In {\em Proceedings of the 17th Conference in Uncertainty in
  Artificial Intelligence ({UAI})}, pages 194--202, 2001.

\bibitem{heckerman:05}
D.~Heckerman.
\newblock A tutorial on learning with bayesian networks.
\newblock Technical Report MSR-TR-95-06, Microsoft Research, March 1995.

\bibitem{DBLP:journals/iandc/Herrmann95}
C.~Herrmann.
\newblock On the undecidability of implications between embedded multivalued
  database dependencies.
\newblock {\em Inf. Comput.}, 122(2):221--235, 1995.

\bibitem{Heckman97}
J.~J. Heckman,  H.~Ichimura and P.~E.~Todd
\newblock Matching as an econometric evaluation estimator: Evidence from evaluating a job training programme.
\newblock {\em The review of economic studies}, 1997, 64(4):605-654.

Ho, Daniel, Kosuke Imai, Gary King, and Elizabeth Stuart. 2005. ‘‘Matching as Nonparametric Preprocessing
for Parametric Causal Inference.’’ http://gking.harvard.edu/files/matchp.pdf.


\bibitem{ho2005}
D.~E.~HO, K~Imai, G.~King and E.~Stuart.
\newblock Matching as Nonparametric Preprocessing
for Parametric Causal Inference.
\newblock \url{http://gking.harvard.edu/files/matchp.pdf}.


Ho, Daniel E. "MatchIt: nonparametric preprocessing for parametric causal inference." PhD diss., Departments of Mental Health and Biostatistics, Johns Hopkins Bloomberg School of Public Health, 1737.

\bibitem{Holland1986}
P.~W. Holland.
\newblock Statistics and causal inference.
\newblock {\em Journal of the American Statistical Association}, 81(396):pp.
  945--960, 1986.

\bibitem{Khoussainova+2012}
N.~Khoussainova, M.~Balazinska, and D.~Suciu.
\newblock Perfxplain: debugging mapreduce job performance.
\newblock {\em Proc. VLDB Endow. (PVLDB)}, 5(7):598--609, Mar. 2012.

\bibitem{kimmig:probprog12}
A.~Kimmig, S.~Bach, M.~Broecheler, B.~Huang, and L.~Getoor.
\newblock A short introduction to probabilistic soft logic.
\newblock In {\em NIPS Workshop on Probabilistic Programming: Foundations and
  Applications}, 2012.

\bibitem{kok:icml05}
S.~Kok and P.~Domingos.
\newblock Learning the structure of {M}arkov logic networks.
\newblock In {\em ICML}, pages 441--448, 2005.

\bibitem{kok:icml09}
S.~Kok and P.~Domingos.
\newblock Learning {M}arkov logic network structure via hypergraph lifting.
\newblock In {\em ICML}, pages 505--512, 2009.

\bibitem{koller:aaai98}
D.~Koller and A.~Pfeffer.
\newblock Probabilistic frame-based systems.
\newblock In {\em Proceedings of the Fifteenth National/Tenth Conference on
  Artificial Intelligence/Innovative Applications of Artificial Intelligence},
  AAAI '98/IAAI '98, pages 580--587, Menlo Park, CA, USA, 1998. American
  Association for Artificial Intelligence.





\bibitem{Lunt14}
M.~Lunt
\newblock Selecting an appropriate caliper can be essential for achieving good balance with propensity score matching.
\newblock In {\em American journal of epidemiology}, 2014, 179(2):226-235.





\bibitem{MaierTOJ10}
M.~E. Maier, B.~J. Taylor, H.~Oktay, and D.~Jensen.
\newblock Learning causal models of relational domains.
\newblock In {\em Proceedings of the Twenty-Fourth {AAAI} Conference on
  Artificial Intelligence, (AAAI)}, 2010.

\bibitem{math-genealogy}
{Mathematical Genealogy}.
\newblock \mbox{\url{http://genealogy.math.ndsu.nodak.edu}}.
\newblock {\em {National Science Foundation}}.


\bibitem{Mitzenmacher05}
M.~Mitzenmacher and E.~Upfal
\newblock Probability and computing: Randomized algorithms and probabilistic analysis.
\newblock  Cambridge University Press, 2005.



\bibitem{DEBulletin2010}
A.~Meliou, W.~Gatterbauer, J.~Halpern, C.~Koch, K.~F. Moore, and D.~Suciu.
\newblock Causality in databases.
\newblock {\em IEEE Data Engineering Bulletin}, Sept. 2010.

\bibitem{MeliouGMS2011}
A.~Meliou, W.~Gatterbauer, K.~F. Moore, and D.~Suciu.
\newblock The complexity of causality and responsibility for query answers and
  non-answers.
\newblock {\em {Proc. VLDB Endow. (PVLDB)}}, 4(1):34--45, 2010.

\bibitem{MeliouGNS2011}
A.~Meliou, W.~Gatterbauer, S.~Nath, and D.~Suciu.
\newblock Tracing data errors with view-conditioned causality.
\newblock In {\em Proceedings of the 2011 ACM SIGMOD International Conference
  on Management of Data}, SIGMOD '11, pages 505--516, 2011.

\ignore{
\bibitem{VLDB14-tutorial}
A.~Meliou, S.~Roy, and D.~Suciu.
\newblock {Causality and Explanations in Databases.}
\newblock {\em {Proc. VLDB Endow. (PVLDB)}}, 7(13):1715--1716, 2014.
\newblock
  \mbox{\url{http://cs.duke.edu/~sudeepa/Tutorial-Causality-Expl-VLDB14.pdf}}.}

\bibitem{DBLP:conf/sigmod/MeliouS12}
A.~Meliou and D.~Suciu.
\newblock Tiresias: the database oracle for how-to queries.
\newblock In {\em Proceedings of the {ACM} {SIGMOD} International Conference on
  Management of Data, {SIGMOD} 2012, Scottsdale, AZ, USA, May 20-24, 2012},
  pages 337--348, 2012.

\bibitem{mihalkova:icml07}
L.~Mihalkova and R.~J. Mooney.
\newblock Bottom-up learning of {M}arkov logic network structure.
\newblock In {\em ICML}, pages 625--632, 2007.




\bibitem{DBLP:journals/jcss/MiklauS07}
G.~Miklau and D.~Suciu.
\newblock A formal analysis of information disclosure in data exchange.
\newblock {\em J. Comput. Syst. Sci.}, 73(3):507--534, 2007.

\bibitem{neyman1923}
J.~Neyman.
\newblock {\em {On the Application of Probability Theory to Agricul- tural
  Experiments. Essay on Principles. Section 9}}.
\newblock PhD thesis, Roczniki Nauk Rolniczych Tom X [in Polish], 1923.
\newblock translated in Statistical Science, 5, page 465-480.

\bibitem{DBLP:journals/ai/NiepertGSG13}
M.~Niepert, M.~Gyssens, B.~Sayrafi, and D.~V. Gucht.
\newblock On the conditional independence implication problem: {A}
  lattice-theoretic approach.
\newblock {\em Artif. Intell.}, 202:29--51, 2013.

\bibitem{DBLP:journals/siamcomp/PaigeT87}
R.~Paige and R.~E. Tarjan.
\newblock Three partition refinement algorithms.
\newblock {\em {SIAM} J. Comput.}, 16(6):973--989, 1987.

\bibitem{Pearl09}
J.~Pearl.
\newblock {\em Causal inference in statistics: An overview. .}.
\newblock  Statistics Surveys, 2009, 3:96-146.

\bibitem{PearlBook2000}
J.~Pearl.
\newblock {\em Causality: models, reasoning, and inference}.
\newblock Cambridge University Press, 2000.

\bibitem{PearlBook2000}
J.~Pearl.
\newblock {\em Causality: models, reasoning, and inference}.
\newblock Cambridge University Press, 2000.


\bibitem{pearl2010introduction}
J.~Pearl.
\newblock An introduction to causal inference.
\newblock {\em The international journal of biostatistics}, 6(2), 2010.

\bibitem{DBLP:conf/aaai/PearlV87}
J.~Pearl and T.~Verma.
\newblock The logic of representing dependencies by directed graphs.
\newblock In {\em Proceedings of the 6th National Conference on Artificial
  Intelligence. Seattle, WA, July 1987.}, pages 374--379, 1987.

\bibitem{Spirtes:book01}
C.~G. Peter~Spirtes and R.~Scheines.
\newblock {\em Causation, Prediction and Search}.
\newblock MIT Press, 2001.

\bibitem{pujara:iswc13}
J.~Pujara, H.~Miao, L.~Getoor, and W.~Cohen.
\newblock Knowledge graph identification.
\newblock In {\em International Semantic Web Conference (ISWC)}, 2013.

\bibitem{RattiganMJ11}
M.~J. Rattigan, M.~E. Maier, and D.~Jensen.
\newblock Relational blocking for causal discovery.
\newblock In {\em Proceedings of the Twenty-Fifth {AAAI} Conference on
  Artificial Intelligence, ({AAAI})}, 2011.

\bibitem{rekatsinas:sigmod12}
T.~Rekatsinas, A.~Deshpande, and L.~Getoor.
\newblock Local structure and determinism in probabilistic databases.
\newblock In {\em ACM SIGMOD Conference}, 2012.

\bibitem{richards:aaai92}
B.~L. Richards and R.~J. Mooney.
\newblock {Learning Relations by Pathfinding}.
\newblock In {\em AAAI}, 1992.

\bibitem{RichardsonD:MLN:2006}
M.~Richardson and P.~Domingos.
\newblock Markov logic networks.
\newblock {\em Mach. Learn.}, 62(1-2):107--136, Feb. 2006.

\bibitem{rosenbaum2002observational}
P.~R. Rosenbaum.
\newblock {\em Observational studies}.
\newblock Springer, 2002.


observed covariates Rosenbaum, P. R. (2009). Design Observational Studies. New York: Springer-Verlag.

\bibitem{rosenbaum2009}
P.~R. Rosenbaum.
\newblock Design observational studies.
\newblock New York, Springer-Verlag, 2009.

\bibitem{RosenbaumRubin1983}
P.~R. Rosenbaum and D.~B. Rubin.
\newblock The central role of the propensity score in observational studies for
  causal effects.
\newblock {\em Biometrika}, 1983, 70(1):41-55.

\bibitem{RosenbaumRubin1984}
P.~R. Rosenbaum and D.~B. Rubin.
\newblock Reducing bias in observational studies using subclassification on the propensity score.
\newblock {\em Journal of the American statistical Association}, 1984, 79(387):516-524.



\bibitem{RoyOS15}
S.~Roy, L.~Orr, and D.~Suciu.
\newblock Explaining query answers with explanation-ready databases.
\newblock In {\em {Under submission}}, 2015.
\newblock \mbox{\url{http://www.cs.duke.edu/~sudeepa/ExplReady.pdf}}.

\bibitem{RoyS14}
S.~Roy and D.~Suciu.
\newblock A formal approach to finding explanations for database queries.
\newblock In {\em Proceedings of the 2014 ACM SIGMOD International Conference
  on Management of Data}, SIGMOD '14, pages 1579--1590, 2014.
\newblock
  \mbox{\url{http://www.cs.duke.edu/~sudeepa/sigmod2014-explanation.pdf}}.

\bibitem{rubin1970thesis}
D.~B. Rubin.
\newblock {\em The Use of Matched Sampling and Regression Adjustment in
  Observational Studies}.
\newblock Ph.D. Thesis, Department of Statistics, Harvard University,
  Cambridge, MA, 1970.

\bibitem{Rubin1974}
D.~B. Rubin.
\newblock Estimating causal effects of treatments in randomized and
  nonrandomized studies.
\newblock {\em {Journal of Educational Psychology}}, 66(5):688, 1974.

\bibitem{rubin1986statistics}
D.~B. Rubin.
\newblock Statistics and causal inference: Comment: Which ifs have causal
  answers.
\newblock {\em Journal of the American Statistical Association},
  81(396):961--962, 1986.

\bibitem{Rubin2005}
D.~B. Rubin.
\newblock Causal inference using potential outcomes.
\newblock {\em Journal of the American Statistical Association},
  100(469):322--331, 2005.

\bibitem{rubin2006matched}
D.~B. Rubin.
\newblock {\em Matched sampling for causal effects}.
\newblock Cambridge University Press, 2006.

\bibitem{sen:uai09}
P.~Sen, A.~Deshpande, and L.~Getoor.
\newblock Bisimulation-based approximate lifted inference.
\newblock In {\em Uncertainty in Artificial Intelligence}, 2009.

\bibitem{sen:vldbj09}
P.~Sen, A.~Deshpande, and L.~Getoor.
\newblock Prdb: Managing and exploiting rich correlations in probabilistic
  databases.
\newblock {\em VLDB Journal, special issue on uncertain and probabilistic
  databases}, 2009.

\bibitem{shalizi:smr11}
C.~Shalizi and A.~Thomas.
\newblock Homophily and contagion are generically coundounded in observational
  social network studies.
\newblock {\em Sociological Methods and Research}, 40, 2011.

\bibitem{sinclair:ajps12}
B.~Sinclair, M.~Mc{C}onnell, and D.~Green.
\newblock Detecting spillover effects: Design and analysis of multilevel
  experiments.
\newblock {\em American Journal of Political Science}, 56(4), 2012.

\bibitem{Stephen16}
S.~A.~, Fenner, R.~Gurjar and T.~Thierauf.
\newblock Bipartite perfect matching
is in quasi-nc.
\newblock {CoRR}, 2016, abs/1601.06319.

\bibitem{studeny:2003}
M.~Studen\'y.
\newblock {\em Probabilistic COnditional Independence Structures}.
\newblock Springer, 2003.

\bibitem{probdbbook}
D.~Suciu, D.~Olteanu, R.~Christopher, and C.~Koch.
\newblock {\em Probabilistic Databases}.
\newblock Morgan \& Claypool Publishers, 1st edition, 2011.

\bibitem{tchetgen:smmr12}
E.~J.~T. Tchetgen and T.~J. Vander{W}eele.
\newblock On causal inference in the presence of interference.
\newblock {\em Statistical Methods Medical Research}, 21(1), 2012.

\bibitem{DBLP:conf/icml/ToulisK13}
P.~Toulis and E.~K. Kao.
\newblock Estimation of causal peer influence effects.
\newblock In {\em Proceedings of the 30th International Conference on Machine
  Learning, {ICML} 2013, Atlanta, GA, USA, 16-21 June 2013}, pages 1489--1497,
  2013.

\bibitem{udrea:sigmod07}
O.~Udrea, L.~Getoor, and R.~Miller.
\newblock Leveraging data and structure in ontology integration.
\newblock In {\em Proceedings of ACM-SIGMOD 2007 International Conference on
  Management}, pages 449--460, 2007.

\bibitem{DBLP:books/cs/Ullman89}
J.~D. Ullman.
\newblock {\em Principles of Database and Knowledge-Base Systems, Volume {II}}.
\newblock Computer Science Press, 1989.

\bibitem{DBLP:conf/stoc/Vardi82}
M.~Y. Vardi.
\newblock The complexity of relational query languages (extended abstract).
\newblock In {\em Proceedings of the 14th Annual {ACM} Symposium on Theory of
  Computing, May 5-7, 1982, San Francisco, California, {USA}}, pages 137--146,
  1982.


\bibitem{Rosenbaum93}
X.~S.~Gu,  P.~R.~Rosenbaum
\newblock Comparison of Multivariate Matching Methods: Structures,
Distances, and Algorithms.
\newblock {\em Journal of Computational and Graphical Statistics}, 1993,  2(4):405-420.



\bibitem{Lunceford04}
J.~K.~Lunceford and M.~Davidian.
\newblock Stratification and Weighting via the Propensity Score in Estimation of Causal Treatment Effects: a Comparative Study.
\newblock {\em  Statistics in medicine}, 2004, 23(19):2937-2960.






\bibitem{Richardson13}
T.~S.~Richardson and J.~M.~Robins.
\newblock Single World Intervention Graphs (SWIGs): A Unification of the Counterfactual and Graphical Approaches to Causality.
\newblock {\em  Center for the Statistics and the Social Sciences, University of Washington Series. }, 2013,  Working Paper 128, no. 30.




\bibitem{wiki-apgar}
{Wikipedia article on APGAR Score}.
\newblock \mbox{\url{http://en.wikipedia.org/wiki/apgar_score}}.
\newblock {\em Wikipedia}.

\bibitem{weather}
{Article America's Worst Weather-Delayed Airports}.
\newblock \mbox{\url{ https://weather.com/travel/news/america-worst-weather-delayed-airports}}.
\newblock {\em weather.com }.


\bibitem{wiki-Hume}
{Wikipedia article on David Hume}.
\newblock \url{{https://en.wikipedia.org/wiki/David_Hume}}.

\bibitem{wikipedia:rubin:causality}
Rubin causal model.
\newblock \url{https://en.wikipedia.org/wiki/Rubin_causal_model}.

\bibitem{barometricpressureheadache:article}
Barometric Pressure and Weather Conditions.
\newblock \url{http://www.barometricpressureheadache.com/barometric-pressure-and-weather-conditions}.


\bibitem{Sekhon08}
J.~S.~ Sekhon.
\newblock The Neyman-Rubin Model of Causal Inference and Estimation via Matching Methods.
\newblock {\em The Oxford handbook of political methodology}, 2008,  pp 271-299.


\bibitem{Shalizi13}
C.~Shalizi.
\newblock Advanced data analysis from an elementary point of view.
\newblock Cambridge University Press, 2013.


\bibitem{flightdata}
{Dataset: Airline On-Time Performance}.
\newblock {\mbox{\url{http://www.transtats.bts.gov/}}}.
\newblock {\em {Bureau of Transportation Statistics, United States Department
  of Transportation}}.

\bibitem{Weatherdata}
The Weather Underground API.
\newblock  {\mbox{\url{http://www.wunderground.com/weather/api/}}}

\bibitem{king15}
G.~King. and R.~Nielsen.
\newblock Why Propensity Scores Should Not Be Used for Matching
\newblock Working paper, 2015.


\end{thebibliography} 